Focasting The Trend of Vehicle Related Crime In Calgary In 2025

### 2024-11-29 paper/paper.qmd
Edit paper and fix cross reference

Urban crime is an ongoing problem compromising public confidence, economic stability, and safety. Among the several types of crimes, vehicle-related offenses include theft of vehicles and theft from vehicles remain a major issue in urban areas across the country. These crimes disturb people's life but also burden local governments and law enforcement. Over the years, Calgary, Canada, has seen varying patterns in vehicle-related crimes, which raises issues regarding future developments of these trends. Designing effective preventive plans, maximizing resource allocation, and guaranteeing community safety all depend on an awareness of current trends and future crime count prediction. This study forecasts vehicle-related crime patterns in Calgary up until 2025, therefore addressing these issues.

@@

The analysis reveals a continued decline in vehicle-related crimes in Calgary, with a mean change of -60.4 each year, and notable mean differences of -435.0 between the two categories. Theft from vehicles is anticipated to show cyclical surges in the summer months, indicating possible opportunistic habits. Conversely, vehicle theft exhibits a more stable decline throughout the course of the year. The data indicate that, although overall vehicle-related crimes are down, seasonal surges in certain categories persist as a problem for policymakers and law enforcement to tackle, for example, the crime count of Theft From Vehicle in August is expected larger than January by 158.4.

@@

#Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

@@

The data analysis and visualization is done in R[@citeR] with the following Packages: tidyverse[@tidy], janitor[@janitor], arrow[@arrow], rstanarm[@rstanarm], ggplot2[@ggplot2], dplyr[@dplyr], here, knitr.

@@

The raw data from Open Data has five columns and 75,595 rows, the column names are displayed below:

```{r}
#| echo: false
#| warning: false
raw_data <- read_csv(here("data/01-raw_data/raw_data.csv"))

@@

# Model

@@

Background details and diagnostics are included in [Appendix -@sec-model-details].

@@

The selection of predictors for the Bayesian regression model is grounded in their expected relationship with the response variable, crime_count, and their importance in explaining and forecasting trends in vehicle-related crimes. Below is a detailed justification for each predictor.

@@

# Results

@@

The regression model estimates, summarized in the @tbl-model-summary below, provide insight into how predictors such as year, month, and category influence vehicle-related crime counts in Calgary. The table includes coefficient estimates and their 95% confidence intervals to quantify the effect sizes and uncertainties associated with each predictor.

```{r}
#| label: tbl-model-summary
#| tbl-cap: Regression Estimates for Crime Counts
#| echo: false
#| warning: false
library(modelsummary)
crime_model <- readRDS(here("models/crime_model.rds"))
modelsummary(crime_model, 
             statistic = c("conf.int"),
             output = "markdown",
             title = "Regression Estimates for Crime Counts")

@@

The coefficient for Year is -60.4, indicating that on average, crime counts decrease by approximately 60 incidents per year for the reference category. This finding aligns with the observed overall decline in vehicle-related crime over time.

@@

Monthly coefficients capture seasonal variations in crime counts relative to the reference month January. July and August exhibit higher-than-average crime counts, with estimates of 124.454 and 159.461, respectively, reflecting seasonal spikes during colder months. In contrast, December shows a lower-than-average count, with a negative estimate of -136.723, indicating fewer crimes during this period compared to January.

@@

The coefficient for Theft OF Vehicle is -437.470, showing that crime counts for "Theft OF Vehicle" are consistently lower than those for "Theft FROM Vehicle." This highlights the fundamental difference in frequency between the two categories. This finding reinforces the visual observation that "Theft FROM Vehicle" consistently exhibits higher counts than "Theft OF Vehicle."

@@

The interaction term for Year and Theft OF Vehicle has an estimate of -0.011, suggesting that the declining trend in crime counts over time is slightly stronger for "Theft OF Vehicle" compared to "Theft FROM Vehicle."

@@

# Discussion

>>>>

Urban crime is an ongoing problem compromising public confidence, economic stability, and safety[@un2012enhancing]. Among the several types of crimes, vehicle-related offenses include theft of vehicles and theft from vehicles remain a major issue in urban areas across the country[@rakauskas2009identification]. These crimes disturb people's life but also burden local governments and law enforcement. Over the years, Calgary, Canada, has seen varying patterns in vehicle-related crimes, which raises issues regarding future developments of these trends. Designing effective preventive plans, maximizing resource allocation, and guaranteeing community safety all depend on an awareness of current trends and future crime count prediction. This study forecasts vehicle-related crime patterns in Calgary up until 2025, therefore addressing these issues.

@@

The analysis reveals a continued decline in vehicle-related crimes in Calgary, with a mean change of -60.5 each year, and notable mean differences of -454.8 between the two categories. Theft from vehicles is anticipated to show cyclical surges in the summer months, indicating possible opportunistic habits. Conversely, vehicle theft exhibits a more stable decline throughout the course of the year. The data indicate that, although overall vehicle-related crimes are down, seasonal surges in certain categories persist as a problem for policymakers and law enforcement to tackle, for example, the crime count of Theft From Vehicle in August is expected larger than January by 159.4. Details are in @sec-result.

@@

The remainder of this paper is structured as follows: @sec-data provides an overview of the data. @sec-model provides the modeling approach. @sec-result present the results and discuss the implications, limitations.  @sec-discussion.

@@

The data analysis and visualization is done in R[@citeR] with the following Packages: tidyverse[@tidy], janitor[@janitor], arrow[@arrow], rstanarm[@rstanarm], ggplot2[@ggplot2], dplyr[@dplyr], here[@here], knitr[@knitr], and bayesplot[@bayes].

@@

The raw data from Open Data has five columns and 75,595 rows, the column names are displayed below: @tbl-raw

```{r}
#| label: tbl-raw
#| tbl-cap: Column Names of Raw Data
#| echo: false
#| warning: false
raw_data <- read_csv(here("data/01-raw_data/raw_data.csv"))

@@

# Model {#sec-model}

@@

Background details and diagnostics are included in Appendix @sec-model-detail.

@@

The selection of predictors for the Bayesian regression model is grounded in their expected relationship with the response variable, crime_count, and their importance in explaining and forecasting trends in vehicle-related crimes. Below is a detailed justification for each predictor. More detailed justification and test of model are in @sec-model-detail

@@

# Results {#sec-result}

@@

The regression model estimates, summarized in the @tbl-model-summary below, provide insight into how predictors such as year, month, and category influence vehicle-related crime counts in Calgary. The table includes coefficient estimates and their standard deviation to quantify the effect sizes and uncertainties associated with each predictor.

```{r}
#| label: tbl-model-summary
#| tbl-cap: Regression Estimates for Crime Counts
#| echo: false
#| warning: false
library(modelsummary)
crime_model <- readRDS(here("models/crime_model.rds"))

# Extract the model summary as a matrix
model_summary <- summary(crime_model)

# Convert the summary into a data frame
summary_table <- as.data.frame(model_summary)

# Rename the columns for clarity
colnames(summary_table) <- c("Mean", "SD", "10%", "50%", "90%")

# Round the "Mean" and "SD" columns to 2 decimal places
summary_table$Mean <- round(summary_table$Mean, 1)
summary_table$SD <- round(summary_table$SD, 1)

kable(summary_table %>% select("Mean", "SD"))

@@

The coefficient for Year is -60.5, indicating that on average, crime counts decrease by approximately 60 incidents per year for the reference category. This finding aligns with the observed overall decline in vehicle-related crime over time.

@@

Monthly coefficients capture seasonal variations in crime counts relative to the reference month January. July and August exhibit higher-than-average crime counts, with estimates of 123.8 and 159.4, respectively, reflecting seasonal spikes during colder months. In contrast, December shows a lower-than-average count, with a negative estimate of -134.8, indicating fewer crimes during this period compared to January.

@@

The coefficient for Theft OF Vehicle is -454.8, showing that crime counts for "Theft OF Vehicle" are consistently lower than those for "Theft FROM Vehicle." This highlights the fundamental difference in frequency between the two categories. This finding reinforces the visual observation that "Theft FROM Vehicle" consistently exhibits higher counts than "Theft OF Vehicle."

@@

The interaction term for Year and Theft OF Vehicle has an estimate of 0.0, suggesting that the declining trend in crime counts over time is statistically no difference between "Theft OF Vehicle" and "Theft FROM Vehicle."

@@

# Discussion {#sec-discussion}
